---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Democratic Support Around the World: Dynamic Comparative Data on Public Support for Democracy"
thanks: "Corresponding author: [frederick-solt@uiowa.edu](mailto:frederick-solt@uiowa.edu).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong Cassandra Tai
  affiliation: Pennsylvania State University
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
anonymous: false
abstract: "[0/150 words]"
keywords: "democracy, democratic support, public opinion, political culture, measurement"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: dcpo-dem_support_data.bib
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apalike
citecolor: black
linkcolor: black
endnotes: yes
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 600,
  fig.width=7,
  fig.height = 2.5
)
# If `DCPOtools` is not yet installed:
# remotes::install_github("fsolt/DCPOtools")

library(DCPOtools)
library(DCPO)
library(tidyverse)
library(here)
library(countrycode)
library(patchwork)
library(ggthemes)
library(rsdmx)

set.seed(324)
```

```{r dcpo_input_raw, include=FALSE, cache = TRUE, cache.extra = tools::md5sum(here::here("data-raw", "surveys.csv"))}
survey_file <- list.files(here("data-raw"),
                          full.names = TRUE) %>% 
  str_subset("surveys.*\\.csv")

surveys <- read_csv(survey_file,
                    col_types = "cccccc") 
# %>% 
#   mutate(item = item_con)

dcpo_input_raw <- DCPOtools::dcpo_setup(vars = surveys,
                                        datapath = here("..",
                                                        "data",
                                                        "dcpo_surveys"),
                                        file = here("data",
                                                    "dcpo_input_raw.csv"))
```

```{r summary_stats, cache = TRUE, cache.extra = tools::md5sum(here::here("data-raw", "dcpo_input_raw.csv"))}
dcpo_input_raw <- read_csv(here("data", "dcpo_input_raw.csv"),
                                  col_types = "cdcddcd")

with_min_coverage <- function(x, min_cov) {
  if (!is.na(min_cov)) {
    country <- year <- years <- spanned <- coverage <- NULL

    x <- x %>%
      group_by(country) %>%
      mutate(years = length(unique(year)),
             spanned = length(min(year):max(year)),
             coverage = years/spanned) %>%
      filter(coverage >= min_cov) %>%
      select(-years, -spanned, -coverage) %>%
      ungroup()
  }
  return(x)
}

process_dcpo_input_raw <- function(dcpo_input_raw_df) {
  dcpo_input_raw_df %>% 
    with_min_yrs(2) %>% 
    with_min_cy(5) %>% 
    with_min_coverage(.2) %>% 
    with_min_yrs(2) %>% # double-check after dropping <5 cy
    filter(year >= 1972 & n > 0) %>% 
    group_by(country) %>% 
    mutate(cc_rank = n()) %>% 
    ungroup() %>% 
    arrange(-cc_rank)
}

dcpo_input_raw1 <- dcpo_input_raw %>% 
  filter(!(
    str_detect(survey, "army_wvs") &
      # WVS obs identified as problematic by Claassen
      ((country == "Albania" & year == 1998) |
         (country == "Indonesia" &
            (year == 2001 | year == 2006)) |
         (country == "Iran" & year == 2000) |
         (country == "Pakistan" &
            (year == 1997 | year == 2001)) | # 1996 in Claassen
         (country == "Vietnam" & year == 2001)
      ) |
      (str_detect(item, "strong_wvs") &
         ((country == "Egypt" & year == 2012) |
            (country == "Iran" &
               (year == 2000 | year == 2007)) | # 2005 in Claassen
            (country == "India") |
            (country == "Pakistan" &
               (year == 1997 | year == 2001)) | # 1996 in Claassen
            (country == "Kyrgyzstan" &
               (year == 2003 | year == 2011)) |
            (country == "Romania" &
               (year == 1998 | year == 2005 | year == 2012)) |
            (country == "Vietnam" & year == 2001)
         )) |
      (survey == "pew2002" &
         (country %in% c("Angola", "Bolivia", "Brazil", "China", "Egypt",
         "Guatemala", "Honduras", "India", "Indonesia", "Côte d'Ivoire",
         "Mali", "Pakistan", "Senegal", "Venezuela", "Vietnam"))
         ) |
      (survey == "pew2005" &
         (country %in% c("China", "India", "Morocco", "Pakistan"))
         ) |
      (survey == "pew2007" &
         (country %in% c("Bolivia", "Brazil", "China", "India", "Côte d'Ivoire", "Pakistan", "South Africa", "Venezuela"))
         ) |
      (
        country %in% c(
          "Puerto Rico",
          "Northern Ireland",
          "SrpSka Republic",
          "Hong Kong SAR China"
        )
      )
  )) %>% 
  process_dcpo_input_raw()

n_surveys <- surveys %>% 
  distinct(survey) %>% 
  nrow()

n_items <- dcpo_input_raw1 %>%
  distinct(item) %>% 
  nrow()

n_countries <- dcpo_input_raw1 %>%
  distinct(country) %>% 
  nrow()

n_cy <- dcpo_input_raw1 %>%
  distinct(country, year) %>% 
  nrow() %>% 
  scales::comma()

n_years <- as.integer(summary(dcpo_input_raw1$year)[6]-summary(dcpo_input_raw1$year)[1])

spanned_cy <- dcpo_input_raw %>% 
  group_by(country) %>% 
  summarize(years = max(year) - min(year) + 1) %>% 
  summarize(n = sum(years)) %>% 
  pull(n) %>% 
  scales::comma()

total_cy <- {n_countries * n_years} %>% 
  scales::comma()

year_range <- paste("from",
                    summary(dcpo_input_raw$year)[1],
                    "to",
                    summary(dcpo_input_raw$year)[6])

n_cyi <- dcpo_input_raw1 %>% 
  distinct(country, year, item) %>% 
  nrow() %>% 
  scales::comma()

back_to_numeric <- function(string_number) {
  string_number %>% 
    str_replace(",", "") %>% 
    as.numeric()
}

covered_share_of_spanned <- {back_to_numeric(n_cy)/back_to_numeric(spanned_cy) * 100}
```

In all, we identified `r n_items` such survey items that were asked in no fewer than five country-years in countries surveyed at least twice; these items were drawn from `r n_surveys` different survey datasets.^[
The complete list of survey items is included in online Appendix A.]

Together, the survey items in the source data were asked in `r n_countries` countries in two or more time points over the `r n_years` years from `r year_range`, giving us a total of `r n_cyi` country-year-item observations.
If every country that appears in the source data had been surveyed in every year in the source data, we would have `r total_cy` observations (assuming each survey asked just one of our items), and if these hypothetical surveys included every _item_, we would have `r {n_countries * n_years * n_items} %>% scales::comma()` observations.
With these figures in mind, the actually available data are very, very sparse.
On the other hand, we do have in the source data `r n_cy` country-years in which we have at least _some_ information about the trust in civil servants of the population, that is, some `r round(covered_share_of_spanned)`% of the `r spanned_cy` country-years spanned by the data we collected.
But there can be no denying that the many different survey items employed renders these data incomparable and difficult to use together.

```{r item_and_country_plots, fig.height = 3.5, fig.width=7, fig.cap = "Countries and Years with the Most Observations in the DSD Source Data \\label{item_country_plots}"}
items_plot <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(item) %>%
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(item, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Country-Years\nObserved") +
  ggtitle("Items")

most_common_item <- dcpo_input_raw1 %>% 
  count(item) %>% 
  arrange(-n) %>% 
  slice_head() %>% 
  pull(item)

most_common_cy <- dcpo_input_raw1 %>% 
  filter(item == most_common_item) %>%
  distinct(country, year) %>%
  nrow()

most_common_surveys <- dcpo_input_raw1 %>%
  filter(item == most_common_item) %>%
  distinct(survey) %>%
  pull(survey) %>% 
  str_split(", ") %>% 
  unlist() %>% 
  unique() %>% 
  sort()


countries_plot <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country)) %>% 
  distinct(country, year, item) %>% 
  count(country) %>%
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95, size = 7),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Year-Items\nObserved") +
  ggtitle("Countries")

cby_plot <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Years\nObserved") +
  ggtitle("Countries")


ybc_plot <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>%
  ggplot(aes(year, nn)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        # axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.text.x = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  xlab("Year") +
  ylab("Countries\nObserved") +
  ggtitle("Year")

us_obs <- dcpo_input_raw1 %>% 
  distinct(country, year, item) %>%
  count(country) %>%
  filter(country == "United States") %>%
  pull(n)

most_common_cyi <- dcpo_input_raw1 %>% 
  count(country, year, item) %>% 
  count(country) %>% 
  arrange(-n) %>% 
  slice_head() %>% 
  pull(country)

most_common_cyi_obs <- dcpo_input_raw1 %>%
  filter(country == most_common_cyi) %>%
  distinct(country, year, item) %>%
  nrow()

most_common_cy <- dcpo_input_raw1 %>% 
  count(country, year) %>% 
  count(country) %>% 
    arrange(-n) %>% 
  slice_head() %>% 
  pull(country)

most_common_cy_obs <- dcpo_input_raw1 %>%
  filter(country == most_common_cyi) %>%
  distinct(country, year, item) %>%
  nrow()

others_cyi <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(country) %>%
  arrange(desc(n)) %>%
  slice(2:5) %>%
  pull(country) %>% 
  knitr::combine_words()

countries_cp <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year, item) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  pull(country)

countries_cbyp <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  pull(country)

adding <- setdiff(countries_cbyp, countries_cp) %>% 
  knitr::combine_words()

dropping <- setdiff(countries_cp, countries_cbyp) %>% 
  knitr::combine_words()

y_peak_year <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>% 
  filter(nn == max(nn)) %>% 
  pull(year)

y_peak_nn <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>% 
  filter(nn == max(nn)) %>% 
  pull(nn)

data_poorest <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(country) %>%
  arrange(n) %>%
  filter(n == 2) %>%
  pull(country) %>% 
  knitr::combine_words() %>% 
  paste0("---", ., "---")

wordify_numeral <- function(x) setNames(c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten", "eleven", "twelve", "thirteen", "fourteen", "fifteen", "sixteen", " seventeen", "eighteen", "nineteen"), 1:19)[x]

n_data_poor <- {data_poorest %>%
    str_split(",") %>% 
    first()} %>% 
  length() 

  if(n_data_poor < 20) {
    n_data_poorest <- n_data_poor %>% 
      wordify_numeral()
  } else {
    n_data_poorest <- n_data_poor
    data_poorest <- " "
  }

world_map <- map_data("world") %>% 
  filter(!long > 180)

cby_map <- world_map %>% 
  distinct(region) %>% 
  mutate(country = countrycode::countrycode(region,
                                            "country.name",
                                            "country.name")) %>% 
  filter(!region=="Antarctica") %>% 
  left_join(dcpo_input_raw1 %>% 
              count(country, year) %>% 
              count(country, name = "Years"),
            by = "country") %>% 
  mutate(Years = ifelse(is.na(Years), 0, Years)) %>% 
  ggplot(aes(fill = Years, map_id = region)) +
  geom_map(map = world_map,
           color = "white",
           size = 0.06) +
  coord_map(projection = "mollweide", 
            ylim=c(-80, 90),
            xlim=c(-170, 170)) +
  theme_void() +
  scale_fill_distiller(na.value = "gray90", 
                       palette = "Blues",
                       direction = 1) +
  ggtitle("Years Observed by Country") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(.05,.1),
        legend.justification = c(0,0), 
        legend.direction = "vertical") +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0))

cby_map + (countries_plot/ ybc_plot) + plot_layout(widths = c(4, 1))
```


\pagebreak
```{r obs_by_cy, fig.height = 9, fig.width = 6.5, fig.cap = "Source Data Observations by Country and Year \\label{obs_by_cy}"}
dcpo_input_raw1 %>% 
  mutate(country = str_replace(country, "’", "'")) %>% 
  distinct(country, year, item, cc_rank) %>% 
  group_by(country, year) %>% 
  summarize(n = n(),
            cc_rank = cc_rank) %>% 
  ungroup() %>% 
  distinct() %>% 
  ggplot(aes(x = year, 
             y = forcats::fct_reorder(country, cc_rank),
             fill = n)) + 
  geom_tile() +
  scale_fill_stepsn(colors = rev(hcl.colors(5, "inferno")),
                    n.breaks = 5,
                    show.limits = TRUE,
                    right = FALSE,
                    name = "Observations") +
  labs(x = NULL, y = NULL) +
  scale_x_continuous(breaks=seq(1972, 2020, 4),
                     sec.axis = dup_axis()) +
  scale_y_discrete(position = "right") +
  theme(legend.justification=c(0, 0), 
        legend.position=c(0.01, 0.01),
        axis.text.y  = element_text(size = 7)) 
```


Consider the most frequently asked item in the data we collected, which asks respondents ...
Employed by the Arab Barometer,the Asia Europe Survey, the Asian Barometer, the British Social Attitudes Survey, the Latino Barometer, the East Asian Social Survey, the European Values Survey, the Italian National Election Study, the South Asian Barometer, and the World Values Survey, this question was asked in a total of `r most_common_cy_obs` different country-years.
That this constitutes only `r {most_common_cy_obs*100/(spanned_cy %>% str_replace(",", "") %>% as.numeric())} %>% round()`% of the country-years spanned by our data---and again, this is the _most common_ survey item---again underscores just how sparse and incomparable the available public opinion data is on this topic.

The left panel of Figure&nbsp;\ref{item_country_plots} shows the dozen countries with the highest count of country-year-item observations.
`r most_common_cyi`, with `r most_common_cyi_obs` observations, is far and away the best represented country in the source data, followed by `r others_cyi`.
At the other end of the spectrum, `r n_data_poorest` countries`r data_poorest`have only the minimum two observations required to be included in the source dataset at all.
The upper right panel shows the twelve countries with the most years observed; this group is similar, but with `r adding` joining the list and `r dropping` dropping off.
The bottom panel counts the countries observed in each year and reveals just how few relevant survey items were asked before 1990.
Country coverage reached its peak in `r y_peak_year`, when respondents in `r y_peak_nn` countries were asked items about trust in civil servants.
In the next section, we describe how we make use of all of this sparse and incomparable survey data to generate complete, comparable time-series using a latent variable model.

```{r dcpo_chunk_options}
evaluate <- TRUE
```



\pagebreak
