---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Dimensionality, Conditionality, and Definition: Why [Recent?] Latent Variable Measures of Democratic Support Are Invalid [idk]"
thanks: "Corresponding author: [frederick-solt@uiowa.edu](mailto:frederick-solt@uiowa.edu).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong Cassandra Tai
  affiliation: Pennsylvania State University
- name: Yue Hu
  affiliation: Tsinghua University
- name: Byung-Deuk Woo
  affiliation: Pohang University of Science and Technology
- name: Frederick Solt
  affiliation: University of Iowa
anonymous: false
abstract: "[0/150 words]"
keywords: "democracy, democratic support, public opinion, political culture, measurement"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: dcpo-dem_support_data.bib
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apalike
citecolor: black
linkcolor: black
endnotes: yes
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 600,
  fig.width=7,
  fig.height = 2.5
)
# If `DCPOtools` is not yet installed:
# remotes::install_github("fsolt/DCPOtools")

library(DCPOtools)
library(cmdstanr)
library(tidyverse)
library(here)
library(countrycode)
library(patchwork)
library(ggthemes)
library(rsdmx)

set.seed(324)
```

```{r dcpo_input_raw, include=FALSE, cache = TRUE, cache.extra = tools::md5sum(here::here("data-raw", "surveys_demsupport.csv"))}
survey_file <- list.files(here("data-raw"),
                          full.names = TRUE) %>% 
  str_subset("surveys.*\\.csv")

surveys <- read_csv(survey_file,
                    col_types = "ccccccc")

dcpo_input_raw <- DCPOtools::dcpo_setup(vars = surveys,
                                        datapath = here("..",
                                                        "data",
                                                        "dcpo_surveys"),
                                        file = here("data",
                                                    "dcpo_input_raw.csv"))
```

```{r summary_stats, cache = TRUE, cache.extra = tools::md5sum(here::here("data-raw", "dcpo_input_raw.csv"))}
dcpo_input_raw <- read_csv(here("data", "dcpo_input_raw.csv"),
                                  col_types = "cdcddcd")

with_min_coverage <- function(x, min_cov) {
  if (!is.na(min_cov)) {
    country <- year <- years <- spanned <- coverage <- NULL

    x <- x %>%
      group_by(country) %>%
      mutate(years = length(unique(year)),
             spanned = length(min(year):max(year)),
             coverage = years/spanned) %>%
      filter(coverage >= min_cov) %>%
      select(-years, -spanned, -coverage) %>%
      ungroup()
  }
  return(x)
}

with_max_gap <- function(x, max_gap, edges = TRUE) {
    if (!is.na(max_gap)) {
        country <- yr_obs <- NULL

        c_yrs <- x %>% 
            group_by(country, year) %>% 
            summarize(year = first(year)) %>% 
            mutate(lead_span = ifelse(!is.na(lead(year)),
                                      lead(year) - year - 1,
                                      50),
                   lag_span = ifelse(!is.na(lag(year)),
                                     year - lag(year) - 1,
                                     50),
                   min_span = pmin(lead_span, lag_span),
                   max_span = pmax(lead_span, lag_span),
                   drop = min_span > max_gap & max_span == 50)
        
        x <- x %>% 
          left_join(c_yrs,
                    by = c("country", "year")) %>% 
          filter(!drop) %>% 
          select(-contains("span")) %>% 
          select(-drop)
    }
    return(x)
}

process_dcpo_input_raw <- function(dcpo_input_raw_df) {
  dcpo_input_raw_df %>% 
    with_min_yrs(3) %>% 
    with_min_cy(5) %>% 
    with_min_yrs(3) %>% # double-check after dropping <5 cy
    filter(year >= 1972 & n > 0) %>% 
    group_by(country) %>% 
    mutate(cc_rank = n()) %>% 
    ungroup() %>% 
    arrange(-cc_rank)
}

dcpo_input_raw1 <- dcpo_input_raw %>% 
  filter(!(
    str_detect(survey, "army_wvs") &
      # WVS obs identified as problematic by Claassen
      ((country == "Albania" & year == 1998) |
         (country == "Indonesia" &
            (year == 2001 | year == 2006)) |
         (country == "Iran" & year == 2000) |
         (country == "Pakistan" &
            (year == 1997 | year == 2001)) | # 1996 in Claassen
         (country == "Vietnam" & year == 2001)
      ) |
      (str_detect(item, "strong_wvs") &
         ((country == "Egypt" & year == 2012) |
            (country == "Iran" &
               (year == 2000 | year == 2007)) | # 2005 in Claassen
            (country == "India") |
            (country == "Pakistan" &
               (year == 1997 | year == 2001)) | # 1996 in Claassen
            (country == "Kyrgyzstan" &
               (year == 2003 | year == 2011)) |
            (country == "Romania" &
               (year == 1998 | year == 2005 | year == 2012)) |
            (country == "Vietnam" & year == 2001)
         )) |
      (survey == "pew2002" &
         (country %in% c("Angola", "Bolivia", "Brazil", "China", "Egypt",
         "Guatemala", "Honduras", "India", "Indonesia", "Côte d'Ivoire",
         "Mali", "Pakistan", "Senegal", "Venezuela", "Vietnam"))
         ) |
      (survey == "pew2005" &
         (country %in% c("China", "India", "Morocco", "Pakistan"))
         ) |
      (survey == "pew2007" &
         (country %in% c("Bolivia", "Brazil", "China", "India", "Côte d'Ivoire", "Pakistan", "South Africa", "Venezuela"))
         ) |
      (
        country %in% c(
          "Puerto Rico",
          "Northern Ireland",
          "SrpSka Republic",
          "Hong Kong SAR China"
        )
      )
  )) %>% 
  process_dcpo_input_raw()

n_surveys <- surveys %>% 
  distinct(survey) %>% 
  nrow()

n_items <- dcpo_input_raw1 %>%
  distinct(item) %>% 
  nrow()

n_countries <- dcpo_input_raw1 %>%
  distinct(country) %>% 
  nrow()

n_cy <- dcpo_input_raw1 %>%
  distinct(country, year) %>% 
  nrow() %>% 
  scales::comma()

n_years <- as.integer(summary(dcpo_input_raw1$year)[6]-summary(dcpo_input_raw1$year)[1])

spanned_cy <- dcpo_input_raw1 %>% 
  group_by(country) %>% 
  summarize(years = max(year) - min(year) + 1) %>% 
  summarize(n = sum(years)) %>% 
  pull(n) %>% 
  scales::comma()

total_cy <- {n_countries * n_years} %>% 
  scales::comma()

year_range <- paste("from",
                    summary(dcpo_input_raw$year)[1],
                    "to",
                    summary(dcpo_input_raw$year)[6])

n_cyi <- dcpo_input_raw1 %>% 
  distinct(country, year, item) %>% 
  nrow() %>% 
  scales::comma()

back_to_numeric <- function(string_number) {
  string_number %>% 
    str_replace(",", "") %>% 
    as.numeric()
}

covered_share_of_spanned <- {back_to_numeric(n_cy)/back_to_numeric(spanned_cy) * 100}

most_common_item
```

## The Latent Variable Approach to Measuring Democratic Support 

We build on the data presented in @Claassen2020a and the larger set collected by @TaiHuSolt2022.
In all, we identified `r n_items` survey items that were asked in no fewer than five country-years in countries surveyed at least three times; these items were drawn from `r n_surveys` different survey datasets.^[
The complete list of survey items is included in online Appendix A.]
Together, the survey items in the source data were asked in `r n_countries` countries in three or more time points over the `r n_years` years from `r year_range`, giving us a total of `r n_cyi` country-year-item observations.
These data are sparse, but they do include `r n_cy` country-years for which we have at least _some_ information about the extent of support for democracy, that is, some `r round(covered_share_of_spanned)`% of the `r spanned_cy` country-years spanned by the data we collected.

```{r item_and_country_plots, fig.height = 3.5, fig.width=7, fig.cap = "Countries and Years with the Most Observations in the Source Data \\label{item_country_plots}"}
items_plot <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(item) %>%
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(item, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Country-Years\nObserved") +
  ggtitle("Items")

countries_plot <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country)) %>% 
  distinct(country, year, item) %>% 
  count(country) %>%
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95, size = 7),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Year-Items\nObserved") +
  ggtitle("Countries")

cby_plot <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Years\nObserved") +
  ggtitle("Countries")

ybc_plot <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>%
  ggplot(aes(year, nn)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        # axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.text.x = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  xlab("Year") +
  ylab("Countries\nObserved") +
  ggtitle("Year")

world_map <- map_data("world") %>% 
  filter(!long > 180)

cby_map <- world_map %>% 
  distinct(region) %>% 
  mutate(country = countrycode::countrycode(region,
                                            "country.name",
                                            "country.name")) %>% 
  filter(!region=="Antarctica") %>% 
  left_join(dcpo_input_raw1 %>% 
              count(country, year) %>% 
              count(country, name = "Years"),
            by = "country") %>% 
  mutate(Years = ifelse(is.na(Years), 0, Years)) %>% 
  ggplot(aes(fill = Years, map_id = region)) +
  geom_map(map = world_map,
           color = "white",
           size = 0.06) +
  coord_map(projection = "mollweide", 
            ylim=c(-80, 90),
            xlim=c(-170, 170)) +
  theme_void() +
  scale_fill_distiller(na.value = "gray90", 
                       palette = "Blues",
                       direction = 1) +
  ggtitle("Years Observed by Country") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(.05,.1),
        legend.justification = c(0,0), 
        legend.direction = "vertical") +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0))

cby_map + (countries_plot/ ybc_plot) + plot_layout(widths = c(4, 1))
```

```{r most_common}
most_common_item <- dcpo_input_raw1 %>% 
  count(item) %>% 
  arrange(-n) %>% 
  slice_head() %>% 
  pull(item)

most_common_item_cy <- dcpo_input_raw1 %>% 
  filter(item == most_common_item) %>%
  distinct(country, year) %>%
  nrow()

most_common_item_surveys <- dcpo_input_raw1 %>%
  filter(item == most_common_item) %>%
  distinct(survey) %>%
  pull(survey) %>% 
  str_split(", ") %>% 
  unlist() %>% 
  unique() %>% 
  sort()

top_country_cyi <- dcpo_input_raw1 %>% 
  distinct(country, year, item) %>%
  count(country) %>%
  arrange(-n) %>% 
  slice_head() %>%
  pull(country)

top_country_cyi_obs <- dcpo_input_raw1 %>%
  filter(country == top_country_cyi) %>%
  distinct(country, year, item) %>%
  nrow()

top_country_cy <- dcpo_input_raw1 %>% 
  count(country, year) %>% 
  count(country) %>% 
  arrange(-n) %>% 
  slice_head() %>% 
  pull(country)

top_country_cy_obs <- dcpo_input_raw1 %>%
  filter(country == top_country_cy) %>%
  distinct(country, year) %>%
  nrow()

countries_cp <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year, item) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  pull(country)

countries_cbyp <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  pull(country)

adding <- setdiff(countries_cbyp, countries_cp) %>% 
  knitr::combine_words()

dropping <- setdiff(countries_cp, countries_cbyp) %>% 
  knitr::combine_words()

y_peak_year <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>% 
  filter(nn == max(nn)) %>% 
  pull(year)

y_peak_nn <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>% 
  filter(nn == max(nn)) %>% 
  pull(nn)

data_poorest <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(country) %>%
  arrange(n) %>%
  filter(n == min(n)) %>%
  pull(country) %>% 
  knitr::combine_words() %>% 
  paste0("---", ., "---")

wordify_numeral <- function(x) setNames(c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten", "eleven", "twelve", "thirteen", "fourteen", "fifteen", "sixteen", " seventeen", "eighteen", "nineteen"), 1:19)[x]

n_data_poor <- {data_poorest %>%
    str_split(",") %>% 
    first()} %>% 
  length() 

if (n_data_poor < 20) {
  n_data_poorest <- n_data_poor %>% 
    wordify_numeral()
} else {
  n_data_poorest <- n_data_poor
  data_poorest <- " "
}

```

The most frequently-asked item in the data we collected appears in `r length(most_common_item_surveys)` different survey datasets, covering a total of `r most_common_item_cy` different country-years.
It asks respondents whether they strongly disagree, disagree, agree, or strongly agree with the statement, "Democracy may have its problems but it is still the best form of government."


The left panel of Figure&nbsp;\ref{item_country_plots} presents a choropleth map indicating the number of years observed in the source data for each country.
`r top_country_cy` leads in this regard; it was observed in `r top_country_cy_obs` different years.
Thanks in part to sustained interest in democratic support in the LatinoBarometer and the Americas Barometer, most Latin America countries were surveyed in at least 25 years.
Germany, Poland, Ukraine, and Russia were all surveyed in more than 20 years.
At the other end of the spectrum, `r n_data_poorest` countries have only the minimum three survey years required to be included in the source dataset at all.

The upper right panel of this figure shows the twelve countries with the most year-items in the source data.
It reveals that South Korea enjoys the richest collection of data on support for democracy.
In addition to countries already mentioned, Japan and the Philippines were also surveyed often and with multiple different items.
The lower right panel counts the countries observed in each year and reveals that relatively few countries were surveyed before the mid-1990s.
Country coverage reached its peak in `r y_peak_year`, when respondents in `r y_peak_nn` countries were asked items about support for democracy.
In the next section, we describe how we make use of all of this sparse and incomparable survey data to generate complete, comparable time-series using a latent variable model.

```{r dcpo_chunk_options, cache=TRUE}
evaluate <- TRUE
```

```{r dcpo_input, eval=TRUE, cache=FALSE, include=FALSE, results=FALSE}
dcpo_input <- DCPOtools::format_dcpo(dcpo_input_raw1,
                                     scale_q = most_common_item,
                                     scale_cp = 2)
save(dcpo_input, file = here::here("data", "dcpo_input.rda"))
```

```{r dcpo, eval=TRUE, cache=FALSE, include=FALSE, results=FALSE}
iter <- 1000

dcpo <- cmdstan_model("~/Documents/Projects/DCPO/inst/stan/dcpo.stan")
dcpo_output <- dcpo$sample(
  data = dcpo_input[1:13], 
  max_treedepth = 14,
  adapt_delta = 0.99,
  step_size = 0.005,
  seed = 324, 
  chains = 4, 
  parallel_chains = 4,
  iter_warmup = iter/2,
  iter_sampling = iter/2,
  refresh = iter/50
)
results_path <- here::here(file.path("data", 
                                     iter, 
                                     {str_replace_all(Sys.time(), "[- :]", "") %>%
                                         str_replace("\\d{2}$", "")}))
dir.create(results_path, 
           showWarnings = FALSE, 
           recursive = TRUE)
dcpo_output$save_data_file(dir = results_path,
                           random = FALSE)
dcpo_output$save_output_files(dir = results_path,
                              random = FALSE)
```

```{r dcpo_results}
if (!exists("results_path")) {
  latest <- "202301131531"
  results_path <- here::here("data", "1000", latest)
  
  # Define OSF_PAT in .Renviron: https://docs.ropensci.org/osfr/articles/auth
  if (!file.exists(file.path(results_path, paste0("dcpo-", latest, "-1.csv")))) {
    dir.create(results_path, showWarnings = FALSE, recursive = TRUE)
    osf_retrieve_node("XXXXX") %>% 
      osf_ls_files() %>% 
      filter(name == latest) %>% 
      osf_download(path = here::here("data", "1000"))
  }
}

dcpo_output <- as_cmdstan_fit(here::here(results_path,
                                         list.files(results_path, pattern = "csv$")))


```

```{r dcpo_summary, cache=TRUE}
load(file = here::here("data", "dcpo_input.rda"))
theta_summary <- DCPOtools::summarize_dcpo_results(dcpo_input,
                                                   dcpo_output,
                                                   "theta")

res_cy <- nrow(theta_summary) %>% 
  scales::comma()

res_c <- theta_summary %>% 
  pull(country) %>% 
  unique() %>% 
  length()

save(theta_summary, file = here::here("data","theta_summary.rda"))
```

```{dcpo_results, cache = TRUE}
theta_results <- edr(par = "theta")


```


```{r cs_plot, fig.cap="DSD Scores, Most Recent Available Year \\label{cs_mry}", fig.height=10, fig.width=8}

n_panes <- 2
axis_text_size <- 10

p1_data <- theta_summary %>%
  group_by(country) %>%
  top_n(1, year) %>%
  ungroup() %>%
  arrange(mean) %>%
  transmute(country_year = paste0(country, " (", year, ")") %>% 
              str_replace("’", "'"),
            estimate = mean,
            conf.high = q90,
            conf.low = q10,
            pane = n_panes - (ntile(mean, n_panes) - 1),
            ranked = as.factor(ceiling(row_number())))

p_theta <- ggplot(p1_data,
                  aes(x = estimate, y = ranked)) +
  geom_segment(aes(x = conf.low, xend = conf.high,
                   y = ranked, yend = ranked),
               na.rm = TRUE,
               alpha = .4) +
  geom_point(fill = "black", shape = 21, size = .5, na.rm = TRUE) +
  theme_bw() + theme(legend.position="none",
                     axis.text.x  = element_text(size = axis_text_size,
                                                 angle = 90,
                                                 vjust = .45,
                                                 hjust = .95),
                     axis.text.y  = element_text(size = axis_text_size),
                     axis.title = element_blank(),
                     strip.background = element_blank(), 
                     strip.text = element_blank(),
                     panel.grid.major = element_line(size = .3),
                     panel.grid.minor = element_line(size = .15)) +
  scale_y_discrete(breaks = p1_data$ranked, labels=p1_data$country_year) +
  coord_cartesian(xlim=c(0, 1)) +
  facet_wrap(vars(pane), scales = "free", nrow = 1)


p_theta +
  plot_annotation(caption = "Note: Gray whiskers represent 80% credible intervals.")

bottom5 <- p1_data %>% 
  arrange(ranked) %>% 
  slice(1:5) %>% 
  pull(country_year) %>% 
  str_replace(" \\(.*", "") %>% 
  knitr::combine_words()

```

We estimated the model using the `DCPOtools` package for R [@Solt2020a], running four chains for 1,000 iterations each and discarding the first half as warmup, which left us with 2,000 samples.
The $\hat{R}$ diagnostic had a maximum value of 1.01, indicating that the model converged.
The dispersion parameters of the survey items indicate that all of them load well on the latent variable (see Appendix A).

The result is estimates, in all `r res_cy` country-years spanned by the source data, of support for democracy.
Figure&nbsp;\ref{cs_mry} displays the most recent available DSD score for each of the `r res_c` countries and territories in the dataset.


The latest scores for `r bottom5` have them as the places where the public expresses the lowest support for democracy.

```{r ts_plots, fig.cap="DSD Over Time Within Selected Countries \\label{ts}", fig.height=3.5}
countries <- c("Finland", "China", "Philippines", "Bangladesh",
               "Malaysia", "Turkey", "United Kingdom", "South Korea",
               "Belgium", "Spain", "United States", "Chile", 
               "Ukraine", "Italy", "Argentina", "Mexico")

c_sum <- theta_summary %>% 
  filter(country %in% countries)

ggplot(data = c_sum, aes(x = year, y = mean)) +
  theme_bw() +
  theme(legend.position = "none") +
  coord_cartesian(xlim = c(1985, 2020), ylim = c(0, 1)) +
  labs(x = NULL, y = "DSD Scores") +
  geom_ribbon(data = c_res, aes(ymin = q10, ymax = q90, linetype=NA), alpha = .25) +
  geom_line(data = c_res) +
  facet_wrap(~country, nrow = 2) +
  theme(axis.text.x  = element_text(size=7,
                                    angle = 90,
                                    vjust = .45,
                                    hjust = .95),
        strip.background = element_rect(fill = "white", colour = "white")) +
  plot_annotation(caption = "Note: Countries are ordered by their DSD scores in their most recent\navailable year; gray shading represents 80% credible intervals.")
```

## Democratic Support as a Latent Variable: Validation Tests
1. dispersion of different items
1. meaning:
    - liberal democratic institutions vs strong-man rule
    - emancipative values vs their absence
1. 

```{r convergent_setup, cache = TRUE, cache.extra = tools::md5sum(here::here("data", "pge.rda"))}

internal_dat <- dcpo_input$data %>% 
  filter(item == "polileader4 3 or higher") %>% 
  mutate(prop = 1 - y_r/n_r,
         se = sqrt((prop*(1-prop))/n),
         prop_90 = prop + qnorm(.9)*se,
         prop_10 = prop - qnorm(.9)*se) %>% 
  left_join(theta_results %>% select(-kk, -tt), by = c("country", "year"))

pge_polileader <- theta_results %>% 
  map(. %>% right_join(dcpo_input$data %>% 
                         filter(item == "polileader4 3 or higher") %>% 
                         rowwise() %>% 
                         transmute(country = country,
                                   year = year,
                                   prop = 1 - y_r/n_r,
                                   se = sqrt((prop*(1-prop))/n),
                                   polileader = rnorm(1, mean = prop, sd = se)), 
                       by = c("country", "year")))

internal_cor <- map(pge_polileader, function(x) with(x, cor(pge_mean, polileader))) %>% 
  unlist() %>% 
  mean() %>% 
  round(2) %>% 
  paste0("R = ", .)

internal_label <- tibble(mean = .78, prop = .99, label = internal_cor)

internal_plot <- internal_dat %>% 
  ggplot(aes(x = mean,
             y = prop * 100)) +
  geom_segment(aes(x = `10%`, xend = `90%`,
                   y = prop * 100, yend = prop * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_segment(aes(x = mean, xend = mean,
                   y = prop_90 * 100, yend = prop_10 * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_bw() +
  theme(legend.position="none",
        axis.text  = element_text(size=8),
        axis.title = element_text(size=9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  coord_cartesian(ylim=c(0,100)) +
  scale_x_continuous(breaks=seq(.1, .9, 0.2)) +
  labs(x = "PGE Score", 
       y = '% Agreeing That "On the Whole, Men\nMake Better Political Leaders than Women Do"') + 
  ggtitle("All Country-Years") +
  geom_label(data = internal_label, aes(label = label), size = 3)

## Cross-sectional
internal_poliequal <- dcpo_input$data %>% 
  filter(item == "poliequal3 2 or higher" & year == 2007) %>% 
  mutate(iso2c = countrycode::countrycode(country,
                             origin = "country.name",
                             destination = "iso2c"),
         prop = 1 - y_r/n_r,
         se = sqrt((prop*(1-prop))/n),
         prop_90 = prop + qnorm(.9)*se,
         prop_10 = prop - qnorm(.9)*se) %>% 
  left_join(theta_results %>% select(-kk, -tt), by = c("country", "year"))

pge_poliequal <- pge %>% 
  map(. %>% right_join(dcpo_input$data %>% 
                         filter(item == "poliequal3 2 or higher" & year == 2007) %>% 
                         rowwise() %>% 
                         transmute(country = country,
                                   year = year,
                                   prop = 1 - y_r/n_r,
                                   se = sqrt((prop*(1-prop))/n),
                                   poliequal = rnorm(1, mean = prop, sd = se)), 
                       by = c("country", "year")))

internal_cs_cor <- map(pge_poliequal, 
                       function(x) with(x, cor(pge_mean, poliequal))) %>% 
  unlist() %>% 
  mean() %>% 
  round(2) %>% 
  paste0("R = ", .)

internal_cs_label <- tibble(mean = .7, prop = .73, label = internal_cs_cor)

internal_cs_plot <- internal_poliequal %>% 
  ggplot(aes(x = mean,
             y = prop * 100)) +
  geom_segment(aes(x = `10%`, xend = `90%`,
                   y = prop * 100, yend = prop * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_segment(aes(x = mean, xend = mean,
                   y = prop_90 * 100, yend = prop_10 * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_text(aes(label = iso2c), size = 2) +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_bw() +
  theme(legend.position="none",
        axis.text  = element_text(size=8),
        axis.title = element_text(size=9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  labs(x = "PGE Score", 
       y = '% Opining That "Men Generally Make\nBetter Political Leaders Than Women"') + 
  ggtitle("Pew Global Spring 2007") +
  geom_label(data = internal_cs_label, aes(label = label), size = 3)

# Longitudinal
internal_us <- dcpo_input$data %>% 
  filter(item == "emopoli2 2 or higher") %>% 
  mutate(prop = y_r/n_r,
         se = sqrt((prop*(1-prop))/n),
         prop_90 = prop + qnorm(.9)*se,
         prop_10 = prop - qnorm(.9)*se) %>% 
  right_join(theta_results %>% 
               select(-kk, -tt) %>%
               filter(country == "United States"),
             by = c("country", "year"))

pge_emopoli <- pge %>% 
  map(. %>% right_join(dcpo_input$data %>% 
                         filter(item == "emopoli2 2 or higher") %>% 
                         rowwise() %>% 
                         transmute(country = country,
                                   year = year,
                                   prop = y_r/n_r,
                                   se = sqrt((prop*(1-prop))/n),
                                   emopoli = rnorm(1, mean = prop, sd = se)), 
                       by = c("country", "year")))

internal_us_cor <- map(pge_emopoli, function(x) with(x, cor(pge_mean, emopoli))) %>% 
  unlist() %>% 
  mean() %>% 
  round(2) %>% 
  paste0("R = ", .)

internal_us_label <- tibble(year = 1980, mean = .87, label = internal_us_cor)

internal_us_plot <- internal_us %>% 
  ggplot(aes(x = year,
             y = mean)) +
  geom_line() +
  geom_ribbon(aes(ymin = `10%`,
                  ymax = `90%`,
                  linetype = NA),
               alpha = .2) +
  geom_point(aes(y = prop),
             fill = "black",
             shape = 21,
             size = .5,
             na.rm = TRUE) +
  geom_path(aes(y = prop),
            linetype = 3,
            na.rm = TRUE,
            alpha = .7) +
  geom_segment(aes(x = year, xend = year,
                   y = prop_90, yend = prop_10),
               na.rm = TRUE,
               alpha = .2) +
  annotate("text", x = 1979, y = .7, size = 2,
           label = 'U.S. GSS') +
  annotate("text", x = 2012, y = .60, size = 2,
           label = "PGE") +
  labs(x = "Year", 
       y = "Score") +
  theme_bw() +
  theme(legend.position="none",
        axis.text  = element_text(size=8),
        axis.title = element_text(size=9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ggtitle("United States") +
  geom_label(data = internal_us_label, aes(label = label), size = 3)

```

```{r convergent, cache = TRUE, cache.extra = tools::md5sum(here::here("data", "pge.rda")), fig.cap="Convergent Validation: Correlations Between PGE Scores and Individual PGE Source Data Survey Items \\label{conv_plot}", fig.height=4}

internal_plot + internal_cs_plot + internal_us_plot +
  plot_annotation(caption = "Note: Gray whiskers and shading represent 80% credible intervals.")
```

Such future research, however, depends on the validity of the PGE scores.
Like @Caughey2019 [684-685], we provide evidence of our measure's validity with convergent validation and construct validation.
Convergent validation refers to showing that a measure is empirically associated with alternative indicators of the same concept [@Adcock2001, 540].
Here, we compare PGE scores to responses to individual source-data survey items that were used to generate our estimates, that is, we provide an 'internal' validation test [see, e.g., @Caughey2019, 689; @Solt2020c, 10].
In the left panel of Figure&nbsp;\ref{conv_plot}, we examine the four-point question on political leaders mentioned above, the most common item in the source data across all country-years.
Then, in the center panel, we look at the question that provides the most data-rich cross-section in the source data, which asked whether respondents felt "men generally make better political leaders than women" and was included in Pew Global's Spring 2007 survey.
Finally, in the right panel, to evaluate how well the PGE scores capture change over time, we focus on the item with the largest number of observations for a single country in the source data, which asked respondents to the U.S. General Social Survey whether they agreed or disagreed that "most men are better suited emotionally for politics than are most women."
In every case, the correlations---estimated taking into account the uncertainty in the measures---are in the expected direction and very strong.
<!-- ^[ -->
<!-- The uncertainty in the PGE score and in the percentage in the population who would agree with the item does not substantially affect the correlation with the political leadership question, but failing to account for this uncertainty would overstate the correlation with the Pew item, at $R =$ `r with(internal_poliequal, cor(prop, mean)) %>% round(2)`, and the U.S. GSS item, at $R =$ `r with(internal_us, cor(prop, mean, use = "pairwise.complete.obs")) %>% round(2)`. -->
<!-- We take up the issue of the importance of taking uncertainty into account when working with the PGE data in a subsequent section.] -->


```{r construct_validation_survey, cache = TRUE, cache.extra = tools::md5sum(here::here("data", "pge.rda")), fig.cap="Construct Validation: Correlations Between PGE Scores and Individual 'Balancing' Gender Egalitarianism Survey Items \\label{surv_const_plot}", fig.height=4}
validation_asianb_raw <- "data-raw/validation_items_asianb.csv" %>% 
  read_csv(col_types = "cccccc") %>% 
  DCPOtools::dcpo_setup(datapath = here("..",
                                        "data",
                                        "dcpo_surveys"),
                        file = here("data",
                                    "dcpo_input_raw.csv"))

dcpo_input_validation_asianb <- validation_asianb_raw %>% DCPOtools::format_dcpo(scale_q = "econ_priority4a",
                                                                                 scale_cp = 2)

econ_priority <- dcpo_input_validation_asianb$data %>% 
  filter(item == "econ_priority4a 3 or higher") %>% 
  mutate(iso2c = countrycode::countrycode(country,
                                          origin = "country.name",
                                          destination = "iso2c"),
         prop = 1 - y_r/n_r,
         se = sqrt((prop*(1-prop))/n),
         prop_90 = prop + qnorm(.9)*se,
         prop_10 = prop - qnorm(.9)*se) %>% 
  left_join(theta_results %>% select(-kk, -tt), by = c("country", "year"))

dsd_econ_priority <- theta_results %>% 
  map(. %>% right_join(econ_priority %>% 
                         rowwise() %>% 
                         transmute(country = country,
                                   year = year,
                                   prop = 1 - y_r/n_r,
                                   se = sqrt((prop*(1-prop))/n),
                                   econ_priority = rnorm(1, mean = prop, sd = se)), 
                       by = c("country", "year")))

econ_priority_cor <- map(dsd_econ_priority, function(x) 
  with(x, cor(pge_mean, econ_priority))) %>% 
  unlist() %>% 
  mean() %>% 
  round(2) %>% 
  paste0("R = ", .)

wantch_label <- tibble(mean = .76, prop = .97, label = wantch_cor)

wantch_plot <- wantch %>% 
  ggplot(aes(x = mean,
             y = prop * 100)) +
  geom_segment(aes(x = `10%`, xend = `90%`,
                   y = prop * 100, yend = prop * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_segment(aes(x = mean, xend = mean,
                   y = prop_90 * 100, yend = prop_10 * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_bw() +
  theme(legend.position="none",
        axis.text  = element_text(size=8),
        axis.title = element_text(size=9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  coord_cartesian(ylim=c(0,100)) +
  labs(x = "PGE Score", 
       y = '% Agreeing That "A Job is Alright But What\nMost Women Really Want is a Home and Children"') + 
  ggtitle("All Country-Years") +
  geom_label(data = wantch_label, aes(label = label), size = 3)

# cross-section
familywoman <- dcpo_input_balance$data %>% 
  filter(item == "familywoman4 3 or higher") %>% 
  mutate(iso2c = countrycode::countrycode(country,
                             origin = "country.name",
                             destination = "iso2c"),
         prop = 1 - y_r/n_r,
         se = sqrt((prop*(1-prop))/n),
         prop_90 = prop + qnorm(.9)*se,
         prop_10 = prop - qnorm(.9)*se) %>% 
  left_join(theta_results %>% select(-kk, -tt), by = c("country", "year"))

pge_familywoman <- pge %>% 
  map(. %>% right_join(familywoman %>% 
                         rowwise() %>% 
                         transmute(country = country,
                                   year = year,
                                   prop = 1 - y_r/n_r,
                                   se = sqrt((prop*(1-prop))/n),
                                   familywoman = rnorm(1, mean = prop, sd = se)), 
                       by = c("country", "year")))

familywoman_cor <- map(pge_familywoman, function(x) 
  with(x, cor(pge_mean, familywoman))) %>% 
  unlist() %>% 
  mean() %>% 
  round(2) %>% 
  paste0("R = ", .)

familywoman_label <- tibble(mean = .8, prop = .97, label = familywoman_cor)

familywoman_plot <- familywoman %>% 
  ggplot(aes(x = mean,
             y = prop * 100)) +
  geom_segment(aes(x = `10%`, xend = `90%`,
                   y = prop * 100, yend = prop * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_segment(aes(x = mean, xend = mean,
                   y = prop_90 * 100, yend = prop_10 * 100),
               na.rm = TRUE,
               alpha = .2) +
  geom_text(aes(label = iso2c), size = 2) +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_bw() +
  theme(legend.position="none",
        axis.text  = element_text(size=8),
        axis.title = element_text(size=9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  coord_cartesian(ylim=c(0,100)) +
  labs(x = "PGE Score", 
       y = '% Agreeing That "The Most Important Role of a\nWoman is to Take Care of Her Home and Family"') + 
  ggtitle("Eurobarometer 87.4") +
  geom_label(data = familywoman_label, aes(label = label), size = 3)

# time series
psw_de <- dcpo_input_balance$data %>% 
  filter(item == "preschoolwork4 3 or higher" & country == "Germany") %>% 
  mutate(prop = y_r/n_r,
         se = sqrt((prop*(1-prop))/n),
         prop_90 = prop + qnorm(.9)*se,
         prop_10 = prop - qnorm(.9)*se) %>% 
  right_join(theta_results %>% 
               select(-kk, -tt) %>%
               filter(country == "Germany"),
             by = c("country", "year"))

pge_de <- pge %>% 
  map(. %>% right_join(psw_de %>% 
                         rowwise() %>% 
                         transmute(country = country,
                                   year = year,
                                   prop = y_r/n_r,
                                   se = sqrt((prop*(1-prop))/n),
                                   var = rnorm(1, mean = prop, sd = se)), 
                       by = c("country", "year")))

psw_de_cor <- map(pge_de, function(x) 
  with(x, cor(pge_mean, var, use = "pairwise.complete.obs"))) %>% 
  unlist() %>% 
  mean() %>% 
  round(2) %>% 
  paste0("R = ", .)

psw_de_label <- tibble(year = 1990, mean = .8, label = psw_de_cor)

psw_de_plot <- psw_de %>% 
  ggplot(aes(x = year,
             y = mean)) +
  geom_line() +
  geom_ribbon(aes(ymin = `10%`,
                  ymax = `90%`,
                  linetype = NA),
               alpha = .2) +
  geom_point(aes(y = prop),
             fill = "black",
             shape = 21,
             size = .5,
             na.rm = TRUE) +
  geom_path(aes(y = prop),
            linetype = 3,
            na.rm = TRUE,
            alpha = .7) +
  geom_segment(aes(x = year, xend = year,
                   y = prop_90, yend = prop_10),
               na.rm = TRUE,
               alpha = .2) +
  annotate("text", x = 2010, y = .30, size = 2,
           label = 'ALLBUS, EVS, WVS') +
  annotate("text", x = 1986, y = .66, size = 2,
           label = "PGE") +
  labs(x = "Year", 
       y = "Score") +
  theme_bw() +
  theme(legend.position="none",
        axis.text  = element_text(size=8),
        axis.title = element_text(size=9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ggtitle("Germany") +
  geom_label(data = psw_de_label, aes(label = label), size = 3)


wantch_plot + familywoman_plot + psw_de_plot +
  plot_annotation(caption = "Note: Gray whiskers and shading represent 80% credible intervals.")
```


## Appendices
\pagebreak
```{r obs_by_cy, fig.height = 9, fig.width = 6.5, fig.cap = "Source Data Observations by Country and Year \\label{obs_by_cy}"}
dcpo_input_raw1 %>% 
  mutate(country = str_replace(country, "’", "'")) %>% 
  group_by(country) %>% 
  mutate(years_rank = length(unique(year))) %>% 
  distinct(country, year, item, cc_rank, years_rank) %>% 
  group_by(country, year) %>% 
  summarize(n = n(),
            cc_rank = cc_rank,
            years_rank = years_rank) %>% 
  ungroup() %>% 
  distinct() %>% 
  ggplot(aes(x = year, 
             y = forcats::fct_reorder(country, years_rank),
             fill = n)) + 
  geom_tile() +
  scale_fill_stepsn(colors = rev(hcl.colors(6, "inferno")),
                    n.breaks = 6,
                    show.limits = TRUE,
                    right = FALSE,
                    name = "Observations") +
  labs(x = NULL, y = NULL) +
  scale_x_continuous(breaks=seq(1972, 2020, 4),
                     sec.axis = dup_axis()) +
  scale_y_discrete(position = "right") +
  theme(legend.justification=c(0, 0), 
        legend.position=c(0.01, 0.01),
        axis.text.y  = element_text(size = 6)) 
```

